---
title: "Mixed Models Assignment"
author: "Leah N. Crowley"
date: "2023-09-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Call packages required for assignment: 
  library(ggplot2)
  library(dplyr)
  library(arm)
  library(MASS)
  library(ggfortify)
  library(nlme)
  library(lme4)
  library(lmerTest)
  library(emmeans)
  library(ggtext)
  library(patchwork)

```

#### Shoutout to [this Youtube video](https://www.youtube.com/watch?v=QCqF-2E86r0) for helping explain some vocab and concepts about mixed models. I still have a *lot* to learn here, but this has been a good start. 

#### The data I used for this assignment came from Dr. Haase. We are using it as a lab to learn more about models during our lab meetings, so I figured I would go ahead and give it a shot for this class. It's about birds and was collected in Georgia, Alabama, and Florida. Stay tuned frro my rough first attempts at modeling with it! 

```{r data, include=TRUE}

# Read in data you are using for this assignment: 
  birds <- read.csv("birds.csv")

```
### Very busy figure showing all of my data points: 
```{r Plot bird data, echo=FALSE, message=FALSE, warning=FALSE}
  ggplot(birds, aes(State, Mass, colour = as.factor(Rep), shape=Sex)) + 
    geom_jitter(width =0.15, size=5) +
    ggtitle("Bird Masses in the South")
          
```
```{r Models and AIC, echo=TRUE}

# First model of data (basic, linear):
  birdy.states <- lmer(Mass ~ (1|State/Rep), data = birds)
  summary(birdy.states)
  anova(birdy.states)
  
# ANOVA for linear model of data: 
  anova(lm(Mass ~ State/Rep, data = birds))

# Interested in differences among the different states; use state as a fixed effect: 
  birdy.states.2 <- lmer(Mass ~ State + (1|State/Rep), data = birds)
  summary(birdy.states.2)
  anova(birdy.states.2)
 
# Test to see which model fits the bird data the best:
  anova(birdy.states.2,birdy.states, test = "Chisq")
  # They are incredibly close (AIC of 1997.5 vs 1999.6). 
```
```{r Different, include=TRUE}

# Fitting a generalized linear mixed model with point ID as a random effect (random intercept)
  glmm <- glmer(Presence ~ Elevation + (1|PointID), data = birds, family = binomial(link = "logit"),
                control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
  summary(glmm)

# Fitting a generalized linear mixed model with point ID as a random effect (random intercept) that varies by state (random slope)
  glmm2 <- glmer(Presence ~ Elevation + (1 + State|PointID), data = birds, family = binomial(link = "logit"))
  summary(glmm2)

# Fitting a generalized linear mixed model with point ID as a random effect only
  glmm3 <- glmer(Presence ~ (1|PointID), data = birds, family = binomial(link = "logit"))
  summary(glmm3)

# Compare models:
  AIC(glmm, glmm2, glmm3)
  
```